FITS format for planetary surfaces: definitions, applications and best practices.

### 2018-08-29 ess_marmo.tex
Update on Overleaf.

However, solar system imagery data have been traditionally described using terrestrial based
reference systems and remote sensing formats.
This is particularly true for planetary surface investigations.
Engineers and cartographers working in the first spatial missions were more familiar with
Earth observation techniques and standards than the astronomy one.
Furthermore, the standardization of the spatial references in FITS dates back to the 2000s,
when planetary surface research habits had already been installed.
In small bodies investigations, where astronomers and planetary scientists have worked
together for longer, FITS format is already more popular.

@@

\subsection{MATISSE}
MATISSE\footnote{https://tools.asdc.asi.it/matisse.jsp} (Multi-purpose Advanced Tool
for the Instruments for the Solar System Exploration) \citep{ZINZI2016} is the webtool
designed by the Space Science Data Center of the Italian Space Agency (SSDC-ASI) to
access, visualize and analyze data from solar system exploration missions.
Currently, MATISSE provides access to public data from five different targets (i.e., Mercury,
Venus, Mars, the dwarf planet Ceres and the asteroid Vesta).
MATISSE 2D outputs are formatted as FITS, in equirectangular projection and in planetocentric coordinates.
The planetary FITS standard described in this work has been implemented in MATISSE in order to make the output
file compliant with most FITS readers, especially DS9 and JS9.
The files generated by MATISSE could be also opened with standard GIS software (e.g.,
QGIS, ArcGIS) 
after being provided as input of the fits2vrt Python library\footnote{https://github.com/epn-vespa/fits2vrt},
whose output is a \texttt{.vrt} GDAL Virtual Header pointing to the FITS file with the appropriate
projection.
Thanks to this approach, in the next major release of MATISSE (v2.0, planned to be
online by the end of 2018) JS9 will be the standard reader for
the 2D online output of the tool, thus allowing users to perform advanced analysis in
both 2D and 3D directly inside a web-based environment.

>>>>

However, solar system imagery data have been traditionally described using terrestrial based
geospatial descriptions and remote sensing formats.
This is particularly true for planetary surface investigations.
Engineers and cartographers working in the first spatial missions were more familiar with
Earth observation techniques and standards than the astronomy one.
Furthermore, the standardization of the spatial references in FITS dates back to the 2000s,
when planetary surface research habits had already been installed.
In small bodies investigations, where astronomers and planetary scientists have worked
together for longer, FITS format is already more popular.

@@

\subsection{MATISSE}
MATISSE\footnote{https://tools.ssdc.asi.it/matisse.jsp} (Multi-purpose Advanced Tool
for the Instruments for the Solar System Exploration) \citep{ZINZI2016} is the webtool
designed by the Space Science Data Center of the Italian Space Agency (SSDC-ASI) to
access, visualize and analyze data from solar system exploration missions.
Currently, MATISSE provides access to public data from five different targets (i.e., Mercury,
Venus, Mars, the dwarf planet Ceres and the asteroid Vesta).
MATISSE 2D outputs are formatted as FITS, in equirectangular projection and in planetocentric coordinates.
The planetary FITS standard described in this work has been implemented in MATISSE in order to make the output
file compliant with most FITS readers, especially DS9 and JS9.
The files generated by MATISSE could be also opened with standard GIS software (e.g.,
QGIS, ArcGIS) 
after being provided as input of the fits2vrt Python library\footnote{https://github.com/epn-vespa/fits2vrt},
whose output is a \texttt{.vrt} GDAL Virtual Header pointing to the FITS file with the appropriate
projection.
Thanks to this approach, in the next major release of MATISSE (v2.0, planned to be
online by the end of 2018) JS9 will be the standard reader for
the 2D online output of the tool, thus allowing users to perform advanced analysis in
both 2D and 3D directly inside a web-based environment.
